---
title: "Exploratory analysis of surprisal in Folketingen"
author: "Malte Lau Petersen"
date: "4 september 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(scales)
# library(lme4)
library(brms)
```

```{r load data, message=FALSE, echo = FALSE} 
lda_output = read_csv("data/nov_tra_res.csv", col_types = cols())
lda_output = read_csv("data/tidy_text.csv", col_types = cols()) %>%
    right_join(lda_output, by = "doc_id") %>%
    mutate(Parti = ifelse(Parti == "RV", "R", Parti),
           Parti = ifelse(Parti == "KF", "K", Parti)) %>%
    mutate(Blok = factor(case_when(
        Parti %in% c("DF", "LA", "V", "K") ~ "Blå Blok",
        Parti %in% c("ALT", "EL", "R", "S", "SF") ~ "Rød Blok",
        TRUE ~ "Grønland/Færøerne"
    ), levels = c("Rød Blok", "Grønland/Færøerne", "Blå Blok")))
```



```{r plot settings}

my_theme = theme_classic() +
    theme(legend.position = c(.92,.25),
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.title.align = 0,
          axis.text = element_text(size = 14),
          plot.title = element_text(size = 22, hjust = .5),
          # plot.subtitle = element_text(size = 12, hjust = .3),
          axis.title = element_text(size = 18))
my_guides = guides(fill = guide_colorbar(title.position = "bottom"))

```


How much do members of Folketinget (Danish Parliament) switch topics? How much do these topic switches get picked up by the next speakers? Barron, Huang, Spang, & DeDeo (2018) found a novelty bias during the French revolution where novel speeches were also more resonant in the subsequent discussions. Will we also find a similar bias in an established democracy such as the Danish one?

### Analysis steps
- Scrape folketinget's web-page for pdfs of meeting minutes.
- Extract the text from the pdfs and tag it with time and speaker information
- TODO: also tag each speech with the role of the speaker (ie ordfører, ordstyrer etc.)
- Pre-process the text:
    + Remove stopwords (including names of the other members of Folketinget)
    + Remove punctuation
    + Transform to lowercase
    + TODO: tokenization with dapipe (udpipe for Danish: http://github.com/ITUnlp/dapipe)
- Do topic modelling where each speech turn is a document, and k=100 topics
- Calculate novelty, transience and resonance
- TODO: clustering
- TODO: group by party or block


### Re-creating figure 1 from Barron et al. (2018) with data from Folketinget
```{r plots, fig.height=4.5, fig.width = 4.5, echo = FALSE}

# transience vs novelty
ggplot(lda_output, aes(novelty, transience)) +
    geom_bin2d(bins = 50) +
    scale_fill_viridis_c(trans = "log", breaks = c(1,10,100,1000,10000),
                         labels = trans_format("log10", math_format(expr = 10^.x, format = force))) +
    geom_smooth(method = "lm", colour = "black", linetype = "dashed",  alpha = 0) +
    my_guides + my_theme +
    labs(x = "Novelty", y = "Transience", title = "Transience vs. Novelty", caption = "Data: Speeches in Folketinget 2009-2017")


# resonance vs novelty
ggplot(lda_output, aes(novelty, resonance)) +
    geom_bin2d(bins = 50) +
    scale_fill_viridis_c(trans = "log", breaks = c(1,10,100,1000),
                         labels = trans_format("log10", math_format(expr = 10^.x, format = force))) +
    geom_smooth(method = "lm", colour = "black", linetype = "dashed", alpha = 0) +
    my_guides + my_theme +
    labs(x = "Novelty", y = "Resonance", title = "Resonance vs. Novelty", caption = "Data: Speeches in Folketinget 2009-2017")

# # resonance vs transience
# ggplot(lda_output, aes(transience, resonance)) +
#     geom_bin2d(bins = 50) +
#     scale_fill_viridis_c(trans = "log", breaks = c(1,10,100,1000),
#                          labels = trans_format("log10", math_format(expr = 10^.x, format = force))) +
#     geom_smooth(method = "lm", colour = "black", linetype = "dashed", alpha = 0) +
#     my_guides + my_theme +
#     labs(x = "Transience", y = "Resonance", title = "Resonance vs. Transience", caption = "Data: Speeches in Folketinget 2009-2017")
```


```{r overall model, include = FALSE, eval = FALSE}
# head(lda_output)
trans = brm(bf(transience ~ novelty * Blok + (1+novelty|Parti),
               sigma ~ 1),
            data = lda_output,
            cores = 2, chains = 2, iter = 200)
save("trans", file = "models/transience.Rdata")
```

```{r show model, eval = FALSE}
summary(trans)
```




# Top speakers

```{r per speaker calc, echo = FALSE}
### per speaker measures
per_speaker = lda_output %>%
    group_by(Name) %>%
    summarise(n = n(),
              `z(Novelty)` = mean(z_novelty),
              `z(Transience)` = mean(z_transience),
              `z(Resonance)` = mean(z_resonance)#,
              # d_r = mean(delta_R)
              )

# names(per_speaker)[6] = "\U0394z(Resonance)"


```

### Most novel

High novelty means a large degree of surprise from the past to the present (ie this is very new language). This can be the sign of an outsider (when paired with high transience/low resonance) or an innovator (when low transience/high resonance)
```{r per speaker tables, results = "asis"}
# most novel (different from past)
per_speaker %>%
    arrange(-`z(Novelty)`) %>%
    filter(n >= 10) %>%
    head(10) %>%
    knitr::kable()
```

<!-- ### Least transient -->
<!-- Low transience means a small degree of surprise from the present to the future (ie this speaker affected what came after it a lot) -->

<!-- ```{r per speaker table 2} -->
<!-- # least transience (least different to future) -->
<!-- per_speaker %>% -->
<!--     arrange(`Transience (z)`) %>% -->
<!--     head(10) %>% -->
<!--     knitr::kable() -->
<!-- ``` -->

### Most resonance
Resonance is calculated as `novelty - transience`

Highly resonant speakers were novel and had a large influence at the same time.

```{r per speaker table 3}
# most resonant (novelty - transience)
per_speaker %>%
    arrange(-`z(Resonance)`) %>%
    filter(n >= 10) %>%
    head(10) %>%
    knitr::kable()
```

# Partier
Averages per parti. n is the number of speeches by that parti
```{r per parti calc, echo = FALSE}
### per parti measures
per_parti = lda_output %>%
    group_by(Parti) %>%
    summarise(n = n(),
              `z(Novelty)` = mean(z_novelty),
              `z(Transience)` = mean(z_transience),
              `z(Resonance)` = mean(z_resonance)
              )

per_parti %>%
    arrange(-`z(Resonance)`) %>%
    knitr::kable()

```


```{r per parti plots}

# transience vs novelty
ggplot(lda_output, aes(novelty, transience, colour = Blok)) +
    # geom_bin2d(bins = 50) +
    geom_point(alpha = .5) + 
    facet_wrap(~Parti) +
    # scale_fill_viridis_c(trans = "log", breaks = c(1,10,100,1000,10000),
    #                      labels = trans_format("log10", math_format(expr = 10^.x, format = force))) +
    # geom_smooth(method = "lm", colour = "black", linetype = "dashed",  alpha = 0) +
    labs(x = "Novelty", y = "Transience", title = "Transience vs. Novelty", caption = "Data: Speeches in Folketinget 2009-2017") +
    theme(legend.position = c(.9, .1)) +
    guides(colour = guide_legend(title = element_blank()))

# resonance vs novelty
ggplot(lda_output, aes(novelty, resonance, colour = Blok)) +
    geom_point(alpha = .5) + 
    facet_wrap(~Parti) +
    theme(legend.position = c(.9, .1)) +
    guides(colour = guide_legend(title = element_blank())) +
    labs(x = "Novelty", y = "Resonance", title = "Resonance vs. Novelty", caption = "Data: Speeches in Folketinget 2009-2017")

```



# Reproducing this analysis
- make sure you installed the dependencies
- run `make`


### References
- Barron, A. T., Huang, J., Spang, R. L., & DeDeo, S. (2018). Individuals, institutions, and innovation in the debates of the French Revolution. Proceedings of the National Academy of Sciences, 115(18), 4607-4612.
